\section*{Angabe zu eingesetzten KI-Hilfsmitteln / Declaration on the Use of AI Tools}
Im Rahmen dieser Arbeit wurden die folgenden KI-Hilfsmittel verwendet. 
Es wurden ausschließlich Tools eingesetzt, die mit den in den Allgemeinen Prüfungsbestimmungen (APB) vorgegebenen Regeln im Einklang stehen und der guten wissenschaftlichen Praxis entsprechen. Die Angaben dienen der Transparenz hinsichtlich Art, Zweck und Umfang der Verwendung.

\medskip

\noindent
\textit{During the preparation of this thesis, the following AI tools were used. Only tools that comply with the rules of the Allgemeinen Prüfungsbestimmungen (APB) and conform with academic integrity were used. This declaration aims to ensure transparency regarding the type, purpose, and extent of usage.}

\medskip
\vspace{0.5cm}
% \begin{tabularx}{\textwidth}{|X|X|X|}
\begin{tabularx}{\textwidth}{|>{\raggedright\arraybackslash}X|
		>{\raggedright\arraybackslash}X|
		>{\raggedright\arraybackslash}X|}
	\hline
	\textbf{KI-Hilfsmittel / AI Tool} & \textbf{Verwendungszweck / Purpose of Use} & \textbf{Einsatzbereich / Scope of Use} \\
	\hline
	ChatGPT (GPT-4o, GPT-5) & Linguistic support, formulation, and refinement of pre-written content & Editing and improving clarity and style of self-written text. No generation of scientific results or novel content. \\
	\hline
	GitHub Copilot (GPT-4o, GPT-5, Claude Sonnet 4) & Programming assistance, syntax suggestions, and code snippets & Support for implementation details. Used to accelerate coding, not for conceptual design or evaluation. \\
	\hline
	ChatGPT Codex & Problem-solving support when stuck with specific coding issues. Analysis of existing code and suggestion of potential fixes & Used occasionally for debugging and generating repair suggestions when independent solutions were not successful. Final integration and validation done independently. \\
	\hline
\end{tabularx}
\vspace{1cm}

\noindent
